Refinement for open automata

### 2021-06-18 2021-Refinement/report.tex
Last typos before sending the report

Refinement relations verifies that and implementation satisfies some properties with respect to a specification.
The open pNets model is based around a hierarchical composition of generalised automata.
A properties a refinement can check on them is the non introduction of deadlocks.
A deadlock being an unintended state where no further action is possible.

@@

The contributions of this report are:
\begin{enumerate}
\item The formalisation of composition for open automata without requiring open pNets.
\item The translation of some properties of refinement relations from LTS to open automata.
\item The characterisation of a composition that doesn't introduce deadlocks.
\item The introduciton of several refinement relations for open automata that behave nicely with composition.
\end{enumerate}
This report is organised as follows:
We begin with Section \ref{sec:sota} by describing the state-of-the-art.
Then there are several sections dedicated to define all the interesting objects and properties:
In section \ref{sec:notations} we give some notations used throughout this paper.
In Section \ref{sec:def} we give a clean and up to date definition of open automata.
In Section \ref{sec:comp} we define their composition without requiring the use of open pNets.
In Section \ref{sec:proofelts} we adapt to open automata several standard properties from refinement relations on LTS.
After that Section \ref{sec:prelref} is dedicated to building refinement relations in the easy cases.
Section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.
Finally Section \ref{sec:ccl} recapitulates the contribution and presents the work left.

@@

Some related work on other models than open automata introduce refinement relations.
In a chapter by Bellegarde, Julliand and Kouchnarenko \cite{10.1007/3-540-46428-X_19}, a simulation relation on transition systems is introduced.
This simulation encompass action refinement, is able to deal with silent actions and is compatible with parallel composition.
Here the refinement relation does not consider action refinement as valid but it should be done in future work.
Also they check how LTL properties are preserved or combined using ther refinement which we do not do.
However their model is less expressive: the transition system model is less expressive than open automata and the parallel composition is less expressive than composition on open automata.
In a later report by Kouchnarenko and Lanoix \cite{10.1007/978-3-540-70881-0_26}, the refinement relation they introduce is on LTS (labelled transition systems).
Their relation is additionally prevents deadlock and livelocks.
The compositon is also extended to synchronised composition which is more expressive.
In our work we also deal with deadlocks but not with livelocks since the latter arise only with silent actions.
This work is closer than the previous one to what we do here, still open automata are more expressive than LTS and composition is more general than sychronised composition.

@@

On the deadlock prevention aspect, an article by Dihego, Sampaio and Oliveira \cite{DIHEGO2020110598} present a refinement relation on process algebra (translated to LTS).
This refinement relation is a special case of inheritance and prevents the introduction of deadlocks.
Their refinement and inheritance are quite the opposite of our refinement in terms of new behaviours.
They have channels, interfaces, inputs and outputs, which in the open automata model can be compared to action labels, holes and action data for both inputs and outputs.
Their model have a rich composition as open automata but the introduction of deadlock is already prevented by a well chosen set of composing operations.
Also their composition is slightly different than the one on open automata because they can cause loops by linking two channels of the same process, where in open automata the composition makes an oriented tree.
In their model there is an explicit deadlock and a succesful end where in open automata there are no explicit end.
We define a deadlock as a configuration without possible transition and assume what is a deadlock when comparing the open automata.
To define their refinement and inheritance relation they use trace and failure semantics, which are weaker than (bi)simulations \cite{10.5555/640428.640430} and could break with open automata compositon.

>>>>

Refinement relations verify that and implementation satisfies some properties with respect to a specification.
The open pNets model is based around a hierarchical composition of generalised automata.
A property a refinement can check on them is the non introduction of deadlocks.
A deadlock being an unintended state where no further action is possible.

@@

The contributions of this report are:
\begin{enumerate}
\item The formalisation of composition for open automata without requiring open pNets.
\item The translation of some properties of refinement relations from LTS to open automata.
\item The characterisation of a composition that doesn't introduce deadlocks.
\item The introduction of several refinement relations for open automata that behave nicely with composition.
\end{enumerate}
This report is organised as follows:
We begin with Section \ref{sec:sota} by describing the state-of-the-art.
Then there are several sections dedicated to define all the interesting objects and properties:
In section \ref{sec:notations} we give some notations used throughout this paper.
In Section \ref{sec:def} we give a clean and up to date definition of open automata.
In Section \ref{sec:comp} we define their composition without requiring the use of open pNets.
In Section \ref{sec:proofelts} we adapt to open automata several standard properties from refinement relations on LTS.
After that Section \ref{sec:prelref} is dedicated to building refinement relations in the easy cases.
Section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.
Finally Section \ref{sec:ccl} recapitulates the contribution and presents the work left.

@@

Some related work on other models than open automata introduce refinement relations.
In a chapter by Bellegarde, Julliand and Kouchnarenko \cite{10.1007/3-540-46428-X_19}, a simulation relation on transition systems is introduced.
This simulation encompass action refinement, is able to deal with silent actions and is compatible with parallel composition.
Here the refinement relation does not consider action refinement as valid but it should be done in future work.
Also they check how LTL properties are preserved or combined using ther refinement which we do not do.
However their model is less expressive: the transition system model is less expressive than open automata and the parallel composition is less expressive than composition on open automata.
In a later report by Kouchnarenko and Lanoix \cite{10.1007/978-3-540-70881-0_26}, the refinement relation they introduce is on LTS (labelled transition systems).
Their relation additionally prevents deadlock and livelocks.
The compositon is also extended to synchronised composition which is more expressive.
In our work we also deal with deadlocks but not with livelocks since the latter arise only with silent actions.
This work is closer than the previous one to what we do here, still open automata are more expressive than LTS and composition is more general than sychronised composition.

@@

On the deadlock prevention aspect, an article by Dihego, Sampaio and Oliveira \cite{DIHEGO2020110598} present a refinement relation on process algebra (translated to LTS).
This refinement relation is a special case of inheritance and prevents the introduction of deadlocks.
Their refinement and inheritance are quite the opposite of our refinement in terms of new behaviours.
They have channels, interfaces, inputs and outputs, which in the open automata model can be compared to action labels, holes and action data for both inputs and outputs.
They have a rich composition as open automata but the introduction of deadlock is already prevented by a well chosen set of composing operations.
Also their composition is slightly different than the one on open automata because they can cause loops by linking two channels of the same process, where in open automata the composition makes an oriented tree.
In their model there is an explicit deadlock and a succesful termination where in open automata there are no explicit termination.
We define a deadlock as a configuration without possible transition and assume what is a deadlock when comparing the open automata.
To define their refinement and inheritance relation they use trace and failure semantics, which are weaker than (bi)simulations \cite{10.5555/640428.640430} and could break with open automata compositon.
